--- 
title: "Dealing With Strings in R"
author: "Lingyun Zhang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
link-citations: yes
links-as-notes: true
colorlinks: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preface {-}

I have studied Chapter 14 of the book, *R for Data Science*, by Garrett Grolemund and Hadley Wickham. I put my notes and other relevant stuff here. 

# Knowledge points

- We can use either single or double quotes to create a string. If string contains quotes, then first use single quotes and then use double quotes, e.g. `'Nelson Mandela said, "It always seems impossible until it\'s done."'`

- Use `\` to escape, e.g. `'\\'` gives the backslash.

- Special characters: `'\n'` (a new line); `"\u00b5"` (Greek letter $\mu$). Type `?'"'` into console to find more.

# Useful functions

All the functions that start with `str_` are from `stringr` package.

```{r, warning=FALSE, message=FALSE}
library(stringr)
```

- String length: `str_length(a_vector_of_strings)`

- string concatenation: `str_c(sring_a, sring_b)` or `paste0(string_a, sting_b)`. NB: use `sep` (e.g. `sep = ","`) to control how they are separated.

- Make a vector of strings become one string: `str_c(a_vector_of_strings, collapse = "")`  or `paste0(a_vector_of_string, collapse = "")`

- Take a sub-string: `str_sub(a_srring, start = ?, end = ?)`. Note that the `start` and `end` can take negative values.

- Change to lower- upper- or title-case: `str_to_lower()`, `str_to_upper()`, `str_to_title()`

- Trim white spaces in a string: `str_trim(a_string, side = c('left', 'right', 'both'))`

- Detect matches: `str_detect(a_string, pattern)`

- Get subset: `str_subset(a_vector_of_string, pattern)`

# Regular expressions

Firstly, what is a regular expression? The best way to answer this question is to give a few examples. 

**Example 1:** `abc` is a regular expression, it means a string `abc`. 

**Example 2:** `.` is a regular expression, and it means any a character. So, `.` is a special character in the regular expression system.

**Example 3:** `\` is a regular expression and it means escape. So, `\` is a special character in the regular system. To literally have the symbol `\`, we use string expression `\\\\`.


**Example 4:** `\.` is **defined** as the symbol dot in the regular expression system, and we use string `\\.` to **express** it.


From the above examples, we should have noticed that we must learn these special characters by heart.

special character   | match/meaning   | string for this symbol
:------------------:|:-------:|:---------------------:
`.`                   | any character | `\\.`
`\`                   | escape  | `\\\\`
`^`                   | the start | `\\^`
`$`                   | the end   | `\\$`
`\d`                    | any digit | `\\d`
`\s`                    | any white space | `\\s`
`[abc]`                 | a or b or c |
`[^abc]`                | anything except a, b or c |
`|`                     | logic operator "or" | `\\|`
`?`                     | zero or one  | `\\?`
`*`                     | zero or more |  `\\*`
`+`                      | one or more |  `\\+`
{n}                      | exactly n   |
{n,}                     | n or more   |
{,m}                     | at most m   |
{n,m}                    | between n and m |




# Examples

## Change the first letter to lowercase {-}

```{r, message=FALSE, warning=FALSE}
library(stringr)

fk_string <- c("Apple Pie", "Chocolate Cake")
str_sub(fk_string, 1, 1) <- str_to_lower(str_sub(fk_string, 1, 1)) 
fk_string
```

## Using `noquote()` {-}

```{r, warning=FALSE, message=FALSE}
fk_df <- data.frame(xyz = 1:3,
                    yzx = 2:4,
                    zxy = 3:5)
list(names(fk_df), noquote(names(fk_df)))
```

## Using `str_detect()` and `str_subset()` {-}

```{r, message=FALSE, warning=FALSE}
library(stringr)
corpus <- stringr::words

(three_letter_words <- corpus[str_detect(corpus, "^[a-z][a-z][a-z]$")])

(seven_or_more_letter_words <- corpus[str_detect(corpus, "^.......+$")])

(start_with_y <- corpus[str_detect(corpus, "^y")])
(end_with_x <- corpus[str_detect(corpus, "x$")])
 
(start_with_vowel <- corpus[str_detect(corpus, "^[aeiou]")])
(not_contain_vowel <- corpus[str_detect(corpus, "^[^aeiou]*$")])
 
(end_with_ed_but_not_eed <-
  corpus[str_detect(corpus, "[^e](ed)$")])

(end_with_ing_or_ise <-
  corpus[str_detect(corpus, "(ing)|(ise)$")])

(start_with_3_consonants <-
  corpus[str_detect(corpus, "^[^aeiou]{3}")])

(three_or_more_vowels_in_a_row <-
  corpus[str_detect(corpus, "[aeiou]{3,}")])

(two_or_more_vowel_consonant_pairs_in_a_row <-
  corpus[str_detect(corpus, "([aeiou][^aeiou]){2,}")])

(start_or_end_with_x <-
  str_subset(corpus, "^x|x$"))
 
(start_with_vowel_and_end_with_consonant <-
  str_subset(corpus, "^[aeiou].*[^aeiou]$"))

a_words <-
  str_subset(corpus, "a")
ae_words <-
  str_subset(a_words, "e")
aei_words <- str_subset(ae_words, "i")
(aeio_words <- str_subset(aei_words, "o"))
(aeiou_words <- str_subset(aeio_words, "u"))


```

## Removing a prefix {-}

```{r, warning=FALSE, message=FALSE}
library(tidyverse)

fk_data <- 
  tibble(a_var = c("10   poor 1", 
                   "20  ok 2", 
                   "30  very good 3",
                   "999 NA"),
         x = rnorm(4))

rm_prefix_df <- 
  fk_data %>% 
  mutate(the_prefix = as.numeric(str_extract(a_var, "^\\d+"))) %>% 
  mutate(a_var = str_replace(a_var, "^(\\d+\\s+)", ""))

rm_prefix_df
```


## Removing a suffix {-}

```{r, warning=FALSE, message=FALSE}
library(tidyverse)

fk_data <- 
  tibble(region = c("A region", 
                    "AB region", 
                    "ABC region"),
         x = rnorm(3))

## Method 1
rm_suffix_df <- 
  fk_data %>% 
  mutate(region = str_replace(region, "\\s(region)$", ""))

(rm_suffix_df)

## Method 2
rm_suffix_df <- 
  fk_data %>%
  mutate(region = str_sub(region, 1, nchar(region) - 7))
(rm_suffix_df)
```

## Roving spaces {-}
```{r, warning=FALSE, message=FALSE}
library(stringr)

x <- "   this  is  a    big    space problem.  "
(x <- str_trim(x, side = "both"))
(x <- str_replace_all(x, "\\s+", " "))
```

## Using `str_split_fixed` {-}

```{r, warning=FALSE, message=FALSE}
library(tidyverse)

fk_df <- 
  tibble(string = 
           c("NZTA: New Zealand Transport Agency",
             "PSU: Primary Sampling Unit",
             "NPD: Non Private Dwelling"))

split_matrix <- str_split_fixed(fk_df$string, ":", n = 2)

split_fk_df <- as.tibble(split_matrix)
names(split_fk_df) <- c("name", "description")
split_fk_df <- 
  split_fk_df %>% 
  mutate(description = str_trim(description, side = "left"))

split_fk_df
```

## Using `filter` and `str_detect` {-}

```{r, warning=FALSE, message=FALSE}
library(tidyverse)

df_1 <- 
  data.frame(x = c("x-18", 
                   sample(letters[1:3], 3), 
                   "a-18", 
                   "b-18"), 
             y = rnorm(6))

df_2 <- 
  df_1 %>% 
  filter(str_detect(x, "-"))

df_2
```


